---
title: "FA4-EDA"
author: "Bayquen, Christopher Gilbert A."
date: "2024-02-27"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## **FA4 : MORTALITY BY LATITUDE DATA**


```{r}
#loading the data set and necessary packages
library(ggplot2)
mortality_df <- read.csv("mortality_by_latitude.csv")
mortality_df

```

<center> 

```{r echo=FALSE}

# Create a scatter plot of mortality index against mean average temperature
p <- ggplot(mortality_df, aes(x = temperature, y = mortality_index)) +
  geom_point(color = "blue", alpha = 0.5) +
  labs(title = "Mortality Index vs. Mean Average Temperature",
       x = "Mean Average Temperature",
       y = "Mortality Index") +
  theme_minimal()

# Add linear regression line to check the trend
p + geom_smooth(method = "lm", se = FALSE,color = "red", linetype = "dashed" )

```
</center>

**Try to identify a transformation of one of the variables that will straighten out the relationship, and make a plot of the residuals to check for any remaining patterns.**

<center>

```{r echo=FALSE}
# Create a scatter plot of transformed mortality index against mean average temperature
plot <- ggplot(mortality_df, aes(x = temperature, y = log10(mortality_index))) +
  geom_point(color = "blue", alpha = 0.5) +
  labs(title = "TransformedMortality Index vs. Mean Average Temperature",
       x = "Mean Average Temperature",
       y = "Transformed log10 Mortality Index") +
  theme_minimal()

# Add linear regression line to check the trend
plot + geom_smooth(method = "lm", se = FALSE,color = "red", linetype = "dashed" )
```


```{r echo=FALSE}

mortality_fit <- lm(mortality_df$mortality_index~mortality_df$temperature)

# Create a scatter plot of mortality index against mean average temperature
p <- ggplot(mortality_fit, aes(x = mortality_df$temperature, y = .resid)) +
  geom_point(color = "blue", alpha = 0.5) +
  labs(title = "Mortality Index vs. Mean Average Temperature Residual Plot") +
  theme_minimal()

# Add linear regression line to check the trend
p + geom_smooth(method = "lm", se = FALSE,color = "red", linetype = "dashed" )
```
</center>

> the residual plot does not show any trend or relationship between the two variables; the residuals are scattered randomly around the horizontal dashed line at y = 0, indicating that the model's errors are random and not correlated with mean average temperature.

## **DIAMONDS DATA SET: a plot of log price as a function of carat with a loess smoother**

```{r echo=FALSE}
data(diamonds)
diamonds




```


<center> 

```{r echo=FALSE}
# Subset the dataset to include only the required variables
diamonds_subset <- diamonds %>%
  select(carat, price) %>%
  mutate(log_price = log(price))  # Create a new variable for log price

# Create the plot
ggplot(diamonds_subset, aes(x = carat, y = log_price)) +
  geom_point() +  # Add points
  geom_smooth(method = "loess", se = FALSE) +  # Add loess smoother without standard error
  labs(x = "Carat", y = "Log Price", title = "Log Price vs. Carat with Loess Smoother") +  # Add labels
  theme_minimal()  # Set minimal theme

```

**Trying several values for the span and degree arguments**

*span = 0.1, degree = 1*

```{r echo=FALSE}
# Subset the dataset to include only the required variables
diamonds_subset <- diamonds %>%
  select(carat, price) %>%
  mutate(log_price = log(price))  # Create a new variable for log price

# Create the plot
ggplot(diamonds_subset, aes(x = carat, y = log_price)) +
  geom_point() +  # Add points
  geom_smooth(method = "loess", se = FALSE, span = 0.1) + 
  labs(x = "Carat", y = "Log Price", title = "Log Price vs. Carat with Loess Smoother") +  # Add labels
  theme_minimal()  # Set minimal theme

```
*span = 0.1*

```{r echo=FALSE}
# Subset the dataset to include only the required variables
diamonds_subset <- diamonds %>%
  select(carat, price) %>%
  mutate(log_price = log(price))  # Create a new variable for log price

# Create the plot
ggplot(diamonds_subset, aes(x = carat, y = log_price)) +
  geom_point() +  # Add points
  geom_smooth(method = "loess", se = FALSE, span = 0.5) + 
  labs(x = "Carat", y = "Log Price", title = "Log Price vs. Carat with Loess Smoother") +  # Add labels
  theme_minimal()  # Set minimal theme

```

*span = 0.5*

**Compare the fit of the loess smoother to the fit of the polynomial + step function regression using a plot of the residuals in the two models**

```{R echo=FALSE}
# Fit a polynomial + step function regression
polynomial_fit <- lm(log_price ~ poly(carat, 3) + I(carat^3), data = diamonds_subset)

# Get predicted values and residuals for the polynomial + step function regression
diamonds_subset$polynomial_predicted <- predict(polynomial_fit)
diamonds_subset$polynomial_residuals <- residuals(polynomial_fit)

# Plot residuals for the polynomial + step function regression
ggplot(diamonds_subset, aes(x = carat, y = polynomial_residuals)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(x = "Carat", y = "Residuals",
       title = "Residuals for Polynomial + Step Function Regression") +
  theme_minimal()

# Fit a loess smoother
loess_fit <- loess(log_price ~ carat, data = diamonds_subset, span = 0.5)

# Get predicted values and residuals for the loess smoother
diamonds_subset$loess_predicted <- predict(loess_fit)
diamonds_subset$loess_residuals <- diamonds_subset$log_price - diamonds_subset$loess_predicted

# Plot residuals for the loess smoother
ggplot(diamonds_subset, aes(x = carat, y = loess_residuals)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "blue") +
  labs(x = "Carat", y = "Residuals",
       title = "Residuals for Loess Smoother") +
  theme_minimal()


```
</center>


> The model with residuals that are closer to zero and show less systematic deviation from zero across the range of predictor values is considered more faithful to the data-- this means that the fit of the polynomial + step function regression is more faithful to the data. 